---
title: "Function_trial"
author: "Philippe"
date: "14 November 2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

In this document, we provide function that extract the group to which an animal belong, the composition of this group and the characteristic of all individuals living in this group. Prior to describing each function, we highlight the database errors that may affect the outcome of the function so that the relevant datacheck can be run before hand
```{r setup, include=FALSE}
library(RMySQL)
library(getPass)
library(EloRating)
library(tidyverse)
library(lubridate)
```

We start by establishing a connection with the database and extract all information we need from there
```{r Database connection}
con <- dbConnect(MySQL(), user = 'philippev',password = getPass(),  
                 dbname = 'Moleratdatabase', host = 'Kalahariresearch.org')

#set seed to get always the same randomization
set.seed(123)

#Among all the info that is extracted one coud see hat e ant put in the function directly


#get the table Colony Membership and prepare it so it can be used repeatedly. Regular datacheck must be carried out on this essential query
#Animal with death colony code but no empty date MemberTo 
#Animal with colony code but empty MemberTo
Membership <- con %>%
  dbGetQuery ("SELECT 
    AnimalRef,
    AnimalID,
    DATE(MemberFrom) AS MemberFrom,
    DATE(MemberTo) AS MemberTo,
    MemberShipBetween.ColonyRef,
    MemberDays,
     MemberShipBetween.Colony AS QueriedColony,
    tblColonyCodes.ColonyOrigin
FROM
    MoleratViews_Pending.MemberShipBetween
LEFT JOIN
    Moleratdatabase.tblColonyCodes ON MoleratViews_Pending.MemberShipBetween.ColonyRef = tblColonyCodes.ColonyRef
WHERE MemberShipBetween.ColonyRef <> 120 
AND MemberShipBetween.Colony <> 'Exported_Nigel'") %>% 
  mutate(MemberFrom=ymd(MemberFrom),MemberTo=ymd(MemberTo)) %>% 
  select(AnimalRef,AnimalID,MemberFrom,MemberTo,QueriedColony,ColonyOrigin)

#get the weight from the database 
WeightList<-con %>% 
  dbGetQuery("SELECT * FROM user_philippev.Weight_AnimalID") %>% 
  mutate(WeightDate=ymd(WeightDate)) %>% 
  select(AnimalID,WeightDate,Weight,WeightType)

#get the individual characteristics. Regular datacheck must be carried out on this essential query
ID_Characteristic <-
  con %>% 
  dbGetQuery("SELECT * FROM user_philippev.ID_Characteristic") %>% 
  mutate(BirthDate=ymd(BirthDate),DeathDate=ymd(DeathDate),Mother_FirstLitter=ymd(Mother_FirstLitter),Father_FirstLitter=ymd(Father_FirstLitter)) %>% 
  select(AnimalID,Sex,Wildcaught,WildcaughtQueen,BirthDate,LitterRef,Mother_FirstLitter,Father_FirstLitter,DeathDate)

```

as an example I will take the list of submissive interaction 
```{r submissive interaction}
#as an example I will take the list of submissive interaction 
#the list must have X,y and z
Winner<- SubCall%>% 
  select(Winner,ObsDate) %>% 
  rename(AnimalID=Winner,Date=ObsDate)
nrow(Winner)
names(SubCall)

```



get_colony function:
List argument must contain AnimalID, Date for which we want to return the colony
```{r get colony}

#get_colony gets the colony of the individual on that date. 
#DF1 gets the list of animal we want to extract the colony. It must have a column specyfing the AnimalID and the Date. If a colony entered manually and subjected to error is present it must be included 
#DF2 gets the Membership table from the database. It must have a Date from and a Date to
#get_colony will return two colonies on the same date if the DF1$Date corespond to the date an individual change colony

get_Colony <- function(DF1, DF2) {
  inner_join(DF1 %>% distinct (AnimalID,Date) #one only wants one colony for each day as individual cannot be measure in two colonies simultaneousls
  , DF2, by = "AnimalID") %>%
  filter(Date >= MemberFrom & Date <= MemberTo)%>% 
  select(-c(MemberFrom,MemberTo,AnimalRef))
}

Colony <- get_Colony(
  SubCall%>% 
    select(Winner,ObsDate)%>%
    rename(AnimalID=Winner,Date=ObsDate),Membership)
View(ExctractedColony)


#get_Colony_Duplicate(), Check the combination of AnimalID and date for which there are more than one Queried Colony per individual for a given date
#highlight that some animal may have been observed on the date they changed group and thus will be assigned to two different groups, therefore generating the potential creation of downstream error when querying group composition
#DF1 is the list of Animal and Date 
#DF2 is the Membership
#DF1 gets updated in the function if no colony has been queried out yet
get_Colony_Duplicate <- function (DF1,DF2){
  if (is.null(DF1$QueriedColony)){
    print ("Colony must be extracted")
    DF1<-get_Colony(DF1,DF2)
  } 
    DF1 %>% 
    mutate(ColonyCount=n()) %>% #could be change for summarized
    filter(ColonyCount>1)
}

ColonyDuplicate<-get_Colony_Duplicate(ExctractedColony,Membership)
View(ColonyDuplicate)


#get_Colony_Mismatch() returns all the rows for which there is a mismatch between a dataframe where the Colony has been entered manually and a dataframe where the colony has been queried from the Membership view of the database
#can be the result of a mistake in data entry OR the result of the animal being assigned to a wrong colony because it was observed the date it changed group
#DF1 is the list from which we the have a column for a colony entered manually, a date and an AnimalId
#DF2 is the data frame returned by get_colony()

get_Colony_Mismatch<-function(DF1,DF2){
left_join(DF1,DF2,by=c("AnimalID","Date") %>% 
   filter(Colony != QueriedColony)
}

ColonyMismatch<-get_Colony_Mismatch(SubCall%>% 
    select(Colony,Winner,ObsDate) %>% 
    rename(AnimalID=Winner,Date=ObsDate),ExctractedColony)


## this is the version that queries out the colony in case that was not done and returns an output in term of distinct combination of AnimalId and Date, not in terms of rows
# get_ColonyMismatch<-function(DF1,DF2){
#   if (is.null(DF1$QueriedColony)){
#     print ("Colony must be extracted")
#     DF1<-get_Colony(DF1,DF2)
#   } 
#   DF1 %>% filter(Colony != QueriedColony)
# }


#ColonyExtracted_NoMismatch() returns the row for which there is no mismatch bethween the colony entered manually
#DF1 is the list from which we the have a column for a colony entered manually, a date and an AnimalId
#DF2 is Membership
    
get_Colony_NoMismatch<-function(DF1,DF2){
  inner_join(DF1 %>% distinct (Colony,AnimalID,Date) #one only wants one colony for each day as individual cannot be measure in two colonies simultaneousls
  , DF2, by = "AnimalID") %>%
  filter(Date >= MemberFrom & Date <= MemberTo) %>% 
  filter(Colony == QueriedColony)
}

Colony_NoMismatch<-get_Colony_NoMismatch(SubCall%>% 
    select(Colony,Winner,ObsDate) %>% 
    rename(AnimalID=Winner,Date=ObsDate),Membership) 

  
View(Colony_NoMismatch)


#get_groupcomp(), extract the group composition of the animals which colony has been established 
#DF1 is the dataframe where a date and a colony are provided and that has has been obtained by get_colony() or get_colony_NoMismatch
#DF2 is Membership from the database

get_groupcomp<-function(DF1,DF2){
  inner_join(DF1 %>% 
  distinct(Date,QueriedColony),DF2,by="QueriedColony") %>% 
  filter(Date >= MemberFrom & Date <=MemberTo) %>% 
  select(QueriedColony,ColonyOrigin,AnimalID,Date)
}

GroupComp<-get_groupcomp(Colony_NoMismatch,Membership)
names(GroupComp)
View(GroupComp)#29745 entires 

#get_closestweight() returns the closest weight of all individuals present in a colony at a given time
#DF1 is a dataframe with a column Date and a column AnimalId. will return the closest weight of all group members at a given date if use output get_groupcomp()
#DF2 is the weight extracted from the database
get_Weight<-function(DF1,DF2){
inner_join(DF1 ,DF2, by="AnimalID") %>%
mutate(DayDiff = abs(round(difftime(WeightDate,Date,units="days")))) %>% 
group_by(AnimalID,Date) %>% 
filter(DayDiff == min(DayDiff)) %>%
ungroup() %>% 
group_by(AnimalID,Date,DayDiff) %>% 
summarise(Weight=mean(Weight)) %>% 
ungroup() %>% 
rename(WeightDayDiff=DayDiff) %>% 
arrange(AnimalID,Date)
}

ClosestWeight<-get_Weight(GroupComp %>%  
    distinct(AnimalID,Date),WeightList)


View(ClosestWeight)#29743 animals were in the group that were observed


#get_IDinfo() extract relevant information from all individual present in a group at a given time
#DF1 provides the AnimalID and dates. will return the info of all group members at a given date if use output get_groupcomp()
#DF2 is the individual info extracted from the database
get_IDinfo<-function(DF1,DF2){
inner_join(DF1,DF2) %>% 
###AGE
mutate(Age=round(difftime(Date,BirthDate,units="days")),
  DeathAge=round(difftime(DeathDate,BirthDate,units="days"))) %>% 
### BREEDING STATUS
mutate(BreedingStatus = ifelse((WildcaughtQueen == 1 |
                                    (!is.na(Mother_FirstLitter) & ((Mother_FirstLitter - 90) < Date))|
                                    (!is.na(Father_FirstLitter) & ((Father_FirstLitter - 90) < Date))),
                                 "Breeder",
                                 ifelse(Sex=="M"&& is.na(Age) && ColonyOrigin == "F",
                                        "Undetermined",
                                        "Helper"))) %>% 
    select(AnimalID,Date,Sex,Wildcaught,LitterRef,Age,BreedingStatus,DeathAge)
    
}

IDinfo<-get_IDinfo(GroupComp,ID_Characteristic)
View(IDinfo)

#get_GroupID_Info
#compute all individual ranks and number of competitor within group at a given date 
#DF1 is a dataframe that contains all the IDinfo (including weight) of complete group at given date 

get_info()<-function(DF1){
IDinfo<-get_IDinfo(GroupComp %>% select(AnimalID,Date),ID_Characteristic)
View(IDinfo)#29745 which is two lines more than weight. Suggest there may be a duplicate in ID_Characteristics
names(IDinfo)
### GROUP BY COLONY AND DATE TO RETURN FOR NEXT MUTATE CALLS
### GROUP SIZE, INDIVIDUAL AND GROUP CHARACTERISTICS REGARDLESS OF SEX
  group_by(QueriedColony,ObsDate) %>% 
  mutate(WeightRank=min_rank(desc(Weight))) %>% # Weight rank
  mutate(AgeRank=min_rank(desc(Age))) %>% # Age rank, issue when originally wild animals, what to do?
  mutate(PupNB=sum(Age<1, na.rm= TRUE)) %>% 
  mutate(PupPresence=ifelse(PupNB==0,"No","Yes")) %>% 
  mutate(MinAge=min(Age,na.rm= TRUE)) %>% 
  mutate(GroupSize=n()) %>% 
  mutate(CompNB_5=sapply(Weight, function(x) sum(abs(x-Weight)<5)-1),
         CompNB_10=sapply(Weight, function(x) sum(abs(x-Weight)<10)-1),
         CompNB_15=sapply(Weight, function(x) sum(abs(x-Weight)<15)-1), 
         CompNB_20=sapply(Weight, function(x) sum(abs(x-Weight)<20)-1)) %>% 
  ungroup() %>% 
  ###UNGROUP
  ### GROUP BY SCAN AND SEX FOR FOLLOWING MUTATE CALLS
  group_by(QueriedColony,ObsDate,Sex) %>% 
  mutate(QueueSize=n()) %>% #queue size
  mutate(WeightRank_Queue=min_rank(desc(Weight))) %>% #Weight rank
  mutate(CompNB_5_Queue=sapply(Weight, function(x) sum(abs(x-Weight)<5)-1),
         CompNB_10_Queue=sapply(Weight, function(x) sum(abs(x-Weight)<10)-1),
         CompNB_15_Queue=sapply(Weight, function(x) sum(abs(x-Weight)<15)-1), 
         CompNB_20_Queue=sapply(Weight, function(x) sum(abs(x-Weight)<20)-1)) %>% 
  ungroup() 
  # UNGROUP
  
#list 
#Join Colony and colony of origin
#Join weight 
#join individual characteristic 
  
```

